%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ACKNOWLEDGEMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{\texorpdfstring{\MakeUppercase{ACKNOWLEDGEMENTS}}{ACKNOWLEDGEMENTS}}
\addcontentsline{toc}{chapter}{ACKNOWLEDGEMENTS}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.


There are certainly too many people to thank for a single page.

I would like to first thank my advisor and chair Dr. Claridge for his
support throughout the entire process, serving as both a mentor and role
model. It has been an absolute pleasure being a graduate student under
his direction, and I have learned so much. 

I want to thank Dr. Culp for his leadership and support of the CC-Compass team.
He has always made sure that we have the resources and backing to make
it all go and he is always our biggest promoter. I'm so glad I got to be part
of a project like CC-Compass.

I'd like to thank committee members Dr. Baltazar and Dr. Rasmussen for
their comments and feedback, helping to vastly improve the quality of
this document.

To the entire CC-Compass team, past and present, especially Kevin
Christman and Sebastian Eluvathingal, it's been an amazing 5 years. I
couldn't imagine a better or more enjoyable work environment, and it's
hard to leave. 

I would also like to thank all of those related to Texas A\&M
volleyball. I've gained so many incredible friends, and I'll never forget
all the experiences playing, coaching, and having fun. 

Special thanks to my family, Mom, Dad, Joe, and Val -- I miss you all
every day and more than you know.

And to Whitney, who has shown incredible patience with all of the
inherent uncertainty in the pursuit of a Ph.D. and has provided all the
personal support to make this a reality.

\pagebreak{}
