%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                           SECTION II
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{\texorpdfstring{\MakeUppercase{Literature Review}}{Literature Review}}


A significant amount of research has been completed in the field of
control optimization schemes. Several different approaches including
direct search methods, non-linear programming, genetic algorithms, and
artificial neural networks have been researched. Optimizations have been
completed at all levels of the HVAC system: Zone, AHU, and Plant.  

When discussing optimization, it is important to be clear on what is
meant by \textit{Controls Optimization}. When discussing the subject
with a controls engineer, they will likely be analyzing parameters such
as the settling time, overshoot, and rise time --- dynamic parameters.
In contrast, this research focused on the optimization of the
\textit{steady-state} behavior of the system. 



%    _____      __              _       __     ____        __ 
%   / ___/___  / /_____  ____  (_)___  / /_   / __ \____  / /_
%   \__ \/ _ \/ __/ __ \/ __ \/ / __ \/ __/  / / / / __ \/ __/
%  ___/ /  __/ /_/ /_/ / /_/ / / / / / /_   / /_/ / /_/ / /__ 
% /____/\___/\__/ .___/\____/_/_/ /_/\__/   \____/ .___/\__(_)
%              /_/                              /_/           


\section{Setpoint Optimization}

Since the time constant of AHU dynamics is on the order of minutes and
the time constant of the thermal loads on the building are on the order
of minutes to hours, it is appropriate to focus on the steady state
behavior of AHUs \cite{Bourdouxhe1998}.

Ke and Mumma were among the first to attempt to balance the benefits of
raising the supply air temperature with the penalties associated with
increased fan energy and decreased dehumidification potential
\cite{Ke1997OptimizedSystems}. This paper was limited in the fact that
it made no discussion on actual implementation. The simulated results
showed that optimizing the balance between fan, cooling, and reheat
energy resulted in approximately 6\% energy savings annually versus a
fixed supply air temperature setpoint. They showed that the benefit was
most promising in the more temperate weather when the VAV system was
running above the minimum primary airflow. Ideally, the fan static
pressure pressure setpoint would be optimized in conjunction with the
supply air temperature as well. 

Wang and Song also balanced supply air temperature with fan power and
included economizer control \cite{Wang2012AirCycles}. They found energy
savings reaching up to 90\% under the specific outdoor air conditions
and space loads in their simulation. The universal control sequence they
proposed relied on an outside air temperature sensor, supply air
temperature sensor, and supply air flow sensor. A hindering assumption
was that the terminal units did not have reheat. It is well known that
terminal unit reheat energy can be a significant contributor to the
total energy consumption and the reduction of reheat is often a source
of significant energy and money savings in existing building
commissioning.  

Qin completed a dissertation in 2014 entitled, ``A Data-Driven Approach
for System Approximation and Set Point Optimization, With a Focus in
HVAC Systems'' \cite{Qin_2014_Res_Letters}. The focus of this work was
related to the programming of thermostats in residential homes and did
not cover any buildings in the commercial sector. The control responses
for the rooms were being optimized and not the temperature setpoints of
the rooms.  

Huh and Brandemuehl optimized the setpoint for hot and humid climates
\cite{Huh2008}. The most significant limitation of this work was that it
was heavily focused on a DX unit that would be common to a retail or
supermarket and was also focused on only a hot and humid climate.
Engdahl and Johansson also devoted investigation into the optimization
of the supply air temperature setpoint in a VAV system
\cite{Engdahl2004}, but only studied AHUs with 100\% outside air. 

While the focus of this research is on air handling units, the research
regarding the optimization of central plants should not be ignored
\cite{Ahn2001, Yu2008OptimizingChillers}. At the plant level, Braun has
completed a great deal of research \cite{braun1990, Braun2007RP1252,
Braun2007HybridCooling, Braun2007GeneralControl}. He has investigated
optimizations of various systems, such as ones for night precooling with packaged
rooftop systems \cite{Braun2005}. Crowther and Furlong showed that
advanced optimization controls for chiller-tower condenser-pump systems
could make a 5-8\% improvement in annual energy use
\cite{Crowther2004}. Henze has also been active in
optimizing thermal storage systems \cite{Henze2005} along with
Kintner-Meyer and Emery \cite{KintnerMeyer1995}. Lu et al. also focused
on the optimization of the entire HVAC system, plant, and building, in a
series of papers \cite{LuLu2004, LuLu2005Part1, LuLu2005Part2,
LuLu2005HVACSystemOptimization}. 

United States patents have even been written on the topic of HVAC
optimization. Cascia has a patent on optimal control for a cooling and
heating plant with DDC control \cite{Cascia1999} and Seem has a patent
that describes the strategy to optimally control an air side economizer
\cite{Seem2002}. 

Numerous other researchers have thoroughly covered many aspects of an
optimally operating HVAC system, which is not limited to
\cite{Gruber2014AlternativeBuildings,Cho2009Single-ductOptimization,Nassif2005,Zaheer-uddin2000OptimalBuildings,Wang2000Model-basedAlgorithm,Henze2003EvaluationSystems,Liu2006ExperimentalFoundation,
Zheng1996,
Ning2010Neuro-optimalSystem,Atthajariyakul2004,Cui2004,XuXinhua2009,SunZhongwei2011,Mumma1997EnergyControl,Mossolly-Ghali-Ghaddar_2009_Energy}.
Wang and Ma provided a comprehensive review of supervisory control
research \cite{Wang2008}. The literature is limited in regards to
efficient methods of rapid implementation. This research also presents a
different optimization methodology that focuses on the use BAS trend
data paired with first-principle models. 



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Advanced Computation Techniques %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\input{AdvancedComputationTechniquesAndControls.tex}
\input{AdvancedTechniquesTable.tex}



% \section{AHU Setpoint/Control Strategy Optimization}

% Mumma has also optimized the ventilation control in a variable air volume system \cite{}.


% Mossolly, Ghali, and Ghaddar examined advanced control strategies based on comfort and indoor air quality \cite{}. One proposed strategy explored adjusting the fresh air rate and the supply air temperature, maintaining a temperature setpoint in the zones and assuring indoor air quality. The second strategy also controlled the fresh air rate and supply air temperature but attempting to maintain thermal comfort based on the predicted mean vote of the zone. They found that they saved 30.4\% in operational costs based on simulations of a building in Beriut during the cooling summer season of 4 months with the control strategy based on the predicted mean vote.

% The work by Mossolly is an example of an advanced, optimized control strategy that does not make much mention of the feasibility of implementation. The strategy is limited by the number of weighting factors in the objective function, and the use of sometimes unreliable \(\text{CO}_{2}\) sensors for indication of demand. 


% Many of these approaches have significant limitations making them difficult to implement and scale. Some of the model-based optimizations include parameters unavailable in typical building BAS systems and the uncertainty in these parameters can be significant. Model specification would be necessary for each building setup, requiring significant engineering work. 

% It is desired to eliminate these limitations by developing a method that focuses only on the most robust sensors typical in a CAV/VAV reheat system, uses simple engineering formulas, and requires very little engineering setup. 
% The optimization literature focuses heavily on the methods and less on the implementation. This work would investigate how a centralized, protected, high-level optimized control program could be used and scaled to numerous buildings that each may have different control vendor programs.  

%     ____  ___   _____    ______                                    
%    / __ )/   | / ___/   / ____/___  ____ ___  ____ ___  __  ______ 
%   / __  / /| | \__ \   / /   / __ \/ __ `__ \/ __ `__ \/ / / / __ \
%  / /_/ / ___ |___/ /  / /___/ /_/ / / / / / / / / / / / /_/ / / / /
% /_____/_/  |_/____/   \____/\____/_/ /_/ /_/_/ /_/ /_/\__,_/_/ /_/ 
                                                                                                                               

\section{BAS Communication}
There are many different networking and communication levels and
protocols related to buildings. Kastner, Neugschwandtner, and Soucek et
al. provided a summary of the different systems that exist in buildings
\cite{Kastner2005}. An understanding of the different levels (described
as management, automation, and field by \cite{Kastner2005}) is important
for developing any automated system. The most important open systems
related to building automation include BACnet, LonWorks, EIB/KNX. Other
important and relative standards, protocols, and technologies are shown
in \tableref{} \ref{tab:CommunicationProtocols}.

\begin{table}
\centering
\begin{tabular}{ c }
\toprule

Network Communication Protocols \\
\midrule
\midrule
BACnet\\ 
LonWorks \\
EIB/KNX\\
IP \\
\\
\midrule
Object Access Protocols \\
\midrule
\midrule
Common Object Request Broker Architecture (CORBA) \\
Java Remote Method Invocation (RMI)     \\
Microsoft Distributed Component Object Model (DCOM)  \\
Simple Object Access Protocol (SOAP) \\
\\
\midrule	
Architecture Style \\
\midrule
\midrule
Representational State Transfer (REST) \\
\\
\midrule
Local Area Network Type \\
\midrule
\midrule
Ethernet \\
ARCNET  \\
Master-Slave/Token-Passing (MS/TP) \\
LonTalk  \\
Point-to-Point (PTP)  \\
\\
\midrule
Data Format \\
\midrule
\midrule
Javascript Object Notation (JSON) \\ 
Extensible Markup Language (XML)  \\
\bottomrule
\end{tabular}
\caption{Important standards/technologies in building automation.}
\label{tab:CommunicationProtocols}
\end{table}


Project Haystack is an important open source initiative that is looking
to bring specific naming conventions to building operational data
(project-haystack.org). It accomplishes this at a more abstract level
than BACnet, LonWorks, and EIB/KNX, using lexical rules. Before data
from sensors in buildings can be used effectively, an ``association'' of
the data to an HVAC object (say data being associated with an air
temperature sensor located in a specific air handling unit) needs to
occur. To create an environment where an intelligent building makes
sense, raw numeric data needs to have extensive meta-data to give it
context. This meta-data can be related to engineering units, what piece
of equipment the data belongs to, the hierarchy of equipment and
relationships to one another, and other modifiers. 



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SUMMARY OF LITERATURE SECTION %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of Literature}

Not surprisingly, the goal of any controlled system is to operate
optimally. It is to be expected that there would be a wealth of
literature on the optimization of all different components in a
particular HVAC system. 

Any optimization problem has objectives and constraints. Many different
objective functions and constraints to optimization problems in HVAC
have been proposed. Numerous different optimization techniques and
methods have been implemented and analyzed by different researchers. 

However, there are several limitations and deficiencies in the
optimization literature. Not all authors have focused on the scalability
of different optimization methodologies. Presently it is not feasible to
implement complicated data-mining algorithms on every individual BAS or
controller, partly due to installation time and partly due to the lack
of facilities managers that have the necessary training to understand
the algorithms and that can keep the system running properly. Facility
managers already have difficulties in maintaining traditional and
straightforward HVAC systems. 

Many of the techniques proposed in the literature also are dependent on
sensors or information that are currently unavailable in typical
commercial HVAC systems. For large-scale implementation, a methodology
must function with a minimal number of sensors and remain useful.

For an individual system, there are more than enough well-documented and
effective optimization methods and algorithms. Few buildings in practice
have these optimizations in place because other numerous challenges
have not been fully solved. These challenges include complexity in
implementing with different BAS vendors, training staff members to
understand the logic behind the optimizations, and simply the technician
effort to install and setup implementations.

This research proposes an intuitive optimization method based on
small first-principle models using historical data that is
designed to efficiently scale. The research investigates how to use
current communication protocols with BAS systems and implementable
methodologies to apply a single set of optimization logic and code to
many and varied air handling units. 

\section{Typical Trends Available from Commissioning Projects}

Unfortunately, at the current time, it is uncommon to have all the
sensor data trended for all portions of the HVAC system.  It is rare to
have the capability to sub-meter the energy use of all the individual
components including the fans, cooling coils, and heating coils. This
section investigates what sensors have been typically available from
existing building commissioning and provides suggestions for which
trends are most important for energy-use breakdowns. 

\subsection{Common Case}

The following suggestions are based on the data for over \num{150000}
trends stored in Implementer along with personal experience. Implementer
is a web application developed by the Energy Systems Laboratory that
aids engineers in the collection and analysis of trend data from
buildings.  The statistical results presented use data until April 13, 
2016. The total number of configured single duct AHUs at this time was
846. The number of trends that existed in these 846 AHUs was
11,806, or approximately 14 trends per AHU. Table
\ref{tab:PointBreakdown} gives the percentage breakdown of the types of
points that we have seen in current Implementer projects. Note that due
to special Implementer considerations, the occurrence percentage
should not be evaluated in absolute terms (due to some projects not being
properly set up at all). The relative relationship between trend types is
the more significant result.  


\begin{table}
\caption{Breakdown of points typically available in single duct AHU systems currently in Implementer.}
\label{tab:PointBreakdown}
\small
\begin{tabular}{L{3cm} L{1.2cm}}
    \parbox[b][][b]{2.9cm}{Point Type} & Occurrence \\
\toprule
65\%+                                  &            \\
\midrule
DAT/SAT                                & 83\%       \\
CHW Valve                              & 73\%       \\
Duct Static Pressure                   & 67\%       \\
\midrule
45\% - 65\%                            &            \\
\midrule
Outdoor Air Damper                     & 65\%       \\
Return Air Temp                        & 58\%       \\
DAT/SAT Setpoint                       & 53\%       \\
Return RH                              & 52\%       \\
Fan Status Points                      & 50\%       \\
Mixed Air Temp                         & 48\%       \\
Return Air CO\textsubscript{2}         & 48\%       \\
\midrule
25\%-45\%                              &            \\
\midrule
Filter \(\Delta\)P                     & 35\%       \\
Static Pressure Stpt.                  & 35\%       \\
Outdoor Air Flow                       & 28\%       \\
Return Air Damper                      & 27\%       \\
Occupied Status                        & 25\%       \\
\end{tabular}
\begin{tabular}{L{3.1cm} r}
\midrule
5\% - 25\%         &      \\
\midrule
Space Temp         & 21\% \\
Supply Air Flow    & 15\% \\
Fan Status         & 15\% \\
Fan S/S            & 14\% \\
Outdoor Air Temp   & 13\% \\
Heating Coil Valve & 21\% \\
Modes              & 10\% \\
Fan Power          & 9\%  \\
Preheat Temp       & 9\%  \\
CCLT               & 8\%  \\
Mixed Air Damper   & 7\%  \\
CHW Supply Temp    & 6\%  \\
Return Air Flow    & 6\%  \\
CHW RT             & 6\%  \\
Air Changes        & 5\%  \\
Fan Proofs         & 5\%  \\
Limits             & 5\%  \\
\end{tabular}
\begin{tabular}{L{3.1cm} r}
\midrule
0\% - 5\%             &               \\
\midrule
\# Boxes In Reheat    & 4\%           \\
Misc. Alarms          & 4\%           \\
Supply Fan kW         & 4\%           \\
Economizer Status     & 3\%           \\
Space Humidity        & 3\%           \\
CCLT Setpoint         & 3\%           \\
\% Load               & 2\%           \\
Cool/Heat Coil Flows  & 2\%           \\
Runtimes              & 2\%           \\
Water Pressure        & 1\%           \\
Fan Volts             & 1\%           \\
HW	 Supply Temp      & \textless 1\% \\
HW Return Temp        & \textless 1\% \\
Fan Current           & \textless 1\% \\
Outdoor Air Flow Stpt & \textless 1\% \\
\end{tabular}
\end{table}

If the building in question consumes district chilled and hot water for
thermal HVAC processes, the importance of additional trends is lessened
since it is clear that the chilled water is used to meet the cooling
load of the building and similarly for the hot water. The electricity
use is then left to lights and equipment, along with the fan and pump energy.

However, if natural gas and electricity are the only energy supply types
for the building, additional information coming from trend data becomes
more important for disaggregating the energy end uses. 

It can normally be assumed that monthly utility bills are available.
Having a smaller time interval on the utility data can also aid
significantly in calibration. Daily data can help aid in
discriminating weekday/weekend profiles, while hourly data can aid in
exposing the diurnal cycle of the building. 


\subsubsection{Fan Power}

In all cases, having the fan power trended will aid in the energy use
breakdown, as this immediately provides information regarding this
portion of the electricity use.  In many cases, it is acceptable to
combine lighting energy and the non-HVAC related internal electricity
use. 

The fan power is trended directly on less than 10\% of systems in
Implementer. The next best option in 15\% of the cases is using the
supply flow along with manufacturers' specifications to estimate
fan power. With the design flow values from mechanical drawings,
fractional power curves based on part-load ratio can be used to estimate
the fan energy use. 

%The cooling energy that results from the chiller/HVAC equipment can be
%estimated from the air-side energy balance across the cooling coils
%(assuming that this building is not using individual fan coil units or
%rooftop units). The cooling energy is a function of the supply air flow,
%mixed air temperature, and supply air temperature. It is also a function
%of the mixed air humidity. 


\subsubsection{Cooling Energy} 

At the air handler level, the sensible cooling energy is related to the
supply flow, and the difference between the mixed air temperature, and
the supply air temperature. The supply flows can vary significantly from
one air handler to another, and also under different loadings. 

The mixed air temperature is the lowest priority of the three air-side
parameters for estimating the sensible cooling energy in the AHU. A
reasonable estimate of the outdoor air temperature is available from
local weather stations, in addition to commonly trended outdoor air
temperatures at the site itself. The return air temperature in air
handling units is normally near the space temperature setpoints. It is
known that the mixed air temperature must be between the return and
outdoor air temperature (if the air streams are indeed being mixed).
Without any other information, the safest estimate for the mixed air
temperature would be halfway between these two measurements. 

As an argument for the claim that mixed air temperature is the lowest priority of
the three parameters, the difference between \(T_{oa}\) and 72\(^\circ\)F
for the hourly outdoor air dry-bulb temperature weather data for College
Station during the year of 2015 was calculated. 1/2 of this
difference would be the worst case error estimation. The median of these
half-differences was 4.9\(^\circ\)F, and the mean was 6.4\(^\circ\)F.
These are robust estimates of the upper bounds on the error of the
estimation of mixed air temperature with no guidance other than the assumption
that the mixed air temperature is between the outdoor air temperature and return
air temperature. With any additional information regarding the outdoor
air fraction, the estimate would be even better. 

Determining the latent load across the cooling coil is difficult using
air-side parameters. Humidity sensors are traditionally unreliable, and
two would be necessary to calculate an absolute humidity
difference.

In this sense, metering the water-side parameters would be a more
reliable estimator of the total cooling load, including the latent
effect. However, trended water-side sensors have not been seen in the
five years of Implementer projects.

\subsubsection{Reheat Energy}

Unless the building has some other large hot water end uses, if the
building consumes natural gas, much of this will be directly related to
heating/reheat. Electricity use will then be distributed between lights
and non-HVAC equipment, fans, along with chiller/HVAC equipment. 

If the building is heated using natural gas, the natural gas consumption
can be an adequate indicator of the level of reheat use in the building.
At the current time, trended data from all terminal unit points is
uncommon. The sensors are typically available, however, the number of
terminal units and a large amount of data to be handled are issues that
cause persons to decline to pursue the collection of the terminal unit
data.

Having the supply air temperature and supply air flow is important
because it not only helps fix the parameters for the cooling energy end
use at the AHU but also the parameters for estimating reheat in the
terminal units. If the terminal units are aggregated together, with the
knowledge of the total flow and supply air temperature, along with
potentially having measured natural gas use, the only parameter left to
estimate reheat is the discharge temperature to the zones. 

\subsubsection{Prioritization} \label{sec:Prioritization}
The following sensors, in no particular order, are the most useful with
regards to energy modeling and determining the energy use breakdown. 
\begin{itemize}
\item Supply air temperature -- This parameter is crucial in estimating both the sensible cooling energy and the reheat energy of the system. 
\item Fan power -- Directly returns the energy end use for fans.
\item Supply air flow -- A key parameter in estimating the flow both at the AHU and the total of which is going to the terminal units, affecting both the cooling and reheat energy end uses.
\item Outdoor air flow -- Important in fixing the ventilation load at the AHU.
\item Terminal unit flows -- Helps set the minimum primary air flow parameter, a sensitive parameter affecting the reheat. 
\item Terminal unit discharge temperatures -- Aids the estimation of
    zone reheat. 
\item Space temperature -- Defines the zone temperatures. 
\item Mixed air temperature -- Aids in the determination of the sensible
    and latent cooling load and estimation of the outdoor air fraction.
\item Return air temperature -- Aids in the estimation of the outdoor air fraction or ventilation load.
\item VSD speed, VFD frequency, etc. -- Indicator of the part-load ratio for the equipment, which may be used in the estimation of fan energy. 
\item Preheat Temperature -- Sets this parameter in the model, aiding in the estimation of the heating end uses.
\end{itemize}

Sensors that may be of medium usefulness:

\begin{itemize}
\item Fan Status or occupied/unoccupied status -- Can indicate the run times and schedules of the building.
\item Supply air static pressure -- Unless the precise location of the
    sensor and the overall duct layout is known, it is of little use in
    calculating the fan power. It does indicate whether the
    fan is on or off which is useful for determining AHU schedules.
\item Various Setpoints -- In a well-controlled building, ideally, the
    value of the controlled sensor will be equivalent to the setpoint.
    However, the setpoint trends may not represent reality, especially
    under conditions of control overrides or other faults.
\item Return air relative humidity -- May be an indicator of the level of latent load in the building. If the return air absolute humidity levels are relatively dry, the latent load may be zero or negligible. 
\item Outdoor air temperatures -- In some circumstances, the temperature
    of the ventilation air may differ from the outdoor air temperature in the
    local community, in which the local measurement will be a better
    indicator of the temperature of the air that is entering the AHU
    from the outdoors. However, the local weather station measurements
    are typically much more reliable and trustworthy than the sensors
    maintained at the site, and this needs to be taken into
    consideration.
\end{itemize}

Sensors that are commonly trended and are not useful for estimating the energy use breakdown of a building:
\begin{itemize}
\item Return air CO\(_2\) -- Not directly related to estimating the energy use.
\item Damper commands -- Since duct layout and fluid flow models are not feasible, the damper commands do not provide information towards the calculation of the energy use. 
\item Chilled water valve/hot water valve -- It is the actuator for controlling the supply air temperature, but does not provide information related to energy use.
\end{itemize}





