%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION I
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\pagestyle{plain} % No headers, just page numbers
\pagenumbering{arabic} % Arabic numerals
\setcounter{page}{1}


\chapter{\uppercase {Introduction: The Importance of Research}}


Adjusting the HVAC control sequences in existing building commissioning is a common method to reduce energy consumption. Many of the original control sequences for building equipment are never optimized or adjusted using detailed engineering, due not only to the amount of time and effort that such an analysis may take, but also due to the lack of skill the on-site maintenance staff may have. 

Sensor data from building automation systems is becoming more abundant as computing resources decrease in price and software improves in quality. This wealth of information can be used in an automated process in order to actively optimize the air conditioning system, without detailed input from an engineer. 

This work attempts to leverage commonly available trend data in order to optimize the setpoint value that are typically used in the air handling units of single duct variable air volume systems. Ideally to be considered optimal, the entire air side system needs to be considered as a whole, including fan energy, cooling energy, and reheat energy. While demand based controls can oftentimes significantly reduce energy use for one component of the system, it does not necessarily optimize the whole. 

It is desired to refrain from adjusting the existing control logic or low level electronics in order to accomplish this outcome. This work attempts to acquire trend data, run the necessary methods to determine the optimal setpoints from a separate dedicated system, and then send the information back and \textit{actively} change the air handling unit setpoints in the BAS. In this way, the methodology can scale easily to many different air handlers and buildings, being indifferent to the vendor of the BAS.